Analytic Light Curves in Reflected Light: \\% Phase Curves, Occultations, and Non-Lambertian Scattering

### 2020-05-12 tex/ms.tex
woot; complete draft

\begin{abstract}
    Abstract here.
    %
    \href{https://github.com/rodluger/starrynight}{\color{linkcolor}\faGithub}
\end{abstract}

%
\section{Introduction}
\label{sec:intro}
%
\xxx{A few paragraphs of introduction here.
    Papers to cite: \citet{Russell1906}, \citet{Cowan2009},
    \citet{Kawahara2010}, \citet{Kawahara2011}, \citet{Fujii2012},
    \citet{Farr2018}, \citet{Louden2018}.}

@@

The piecewise nature of the illumination function at the day/night terminator
makes the problem of computing the visible flux particularly difficult; most
studies to date have tackled the problem numerically, by discretizing the surface
and computing the relevant integrals by summing over the visible
pixels \xxx{(cite)}. Recently, \citet{Haggard2018} developed an analytic
framework for computing light curves of unocculted bodies illuminated
by a point source. In this paper, we re-derive their solution under the
\starry framework and extend it for the first time to the case where
the body is occulted by another spherical body, which may or may not be the
illumination source.

@@

@@

The algorithm presented in this paper has been implemented in the \Python
package \starry, which can be installed from
\href{https://github.com/rodluger/starry}{GitHub} or via the Python package
manager \textsf{pip}. The algorithm is coded in a mixture of \textsf{C++}
with forward automatic differentiation using the \textsf{Eigen} library
\citep{eigen} and \Python with backward differentiation using
just-in-time compiled \textsf{theano}
operations \citep{theano}. The user interface, however, is purely in
\Python. The \textsf{theano} backend facilitates integration with the
\textsf{exoplanet} modeling package \citep{exoplanet} and in particular
with \textsf{pymc3} \citep{pymc3} for inference with gradient-based
Markov Chain Monte Carlo (MCMC) schemes.
%
Complete \href{https://rodluger.github.io/starry}{documentation} and an
extensive library of tutorials is available online.
The links next to each
of the figures \codeicon\, point to the \Python scripts used to generate
them and may also help in learning how to use \starry.

@@

The third limitation concerns the flexibility of the \starry model.
While we presented ways to capture non-Lambertian scattering
in \starry (\S\ref{sec:nonlambertian}), there are certain aspects
of light curves in reflected light that cannot be captured by the model.
One example of this is shadowing. Craters on the moon or
volcanoes on Io can cast large shadows visible from space, particularly
if viewed near crescent phase. Unfortunately, there is no way to model
this within the \starry framework. Another example is specular reflection,
or ``glint'', which is a pronounced signal for the Earth due its
oceans \citep{Robinson2014} and on Titan due to
its hydrocarbon lakes \citep{Barnes2011}. In principle, glint could be
modeled in the same way as non-Lambertian sccattering, by constructing the
linear operator $\mathbf{I}$ in such a way as to downweight portions of the
projected disk where $\vartheta_i \neq \vartheta_r$. In practice, however,
if the size of the glint spot is small (which is typically the case),
an expansion at extremely high $l$ ($l \sim 360$ in the case of the Earth)
would be required, which would not work for the reasons above.
Instead, it may be possible to combine the \starry algorithm with the
formalism of \citet{Haggard2018}, who derived analytic expressions for
phase curves of delta function maps ($\delta$-map), to model glint.
As discussed in \citet{LustigYaeger2018}, glint mapping is an extremely
powerful way to not only map terrestrial planets but also to confirm
their habitability via the presence of an ocean.

Finally, we would like to emphasize that while spherical harmonics are
a convenient basis for the purpose of computing light curves, they
have a significant drawback when it comes to modeling real planetary
surfaces: it can be difficult to strictly enforce physical
values of the albedo everywhere on the surface map when doing inference.
That is because there is no analytic way to determine whether a
spherical harmonic representation is positive-valued
(or restricted to a given range) everywhere on the sphere.
Instead, this must be checked numerically, by evaluating the function on a
discrete grid. This makes it somewhat cumbersome to implement positivity as a prior
when doing inference; in particular, this prior cannot be expressed as
a Gaussian, so the analytic expression for the posterior
discussed in \S\ref{sec:linearity} will generally have nonzero support for
negative albedo values. This is particularly problematic when the data is
not very constraining and the posterior is prior-dominated, as
positivity can be an \emph{extremely} informative prior
\citep[e.g.,][]{Fienup1978}.
We therefore recommend that in such cases
users of the \starry algorithm use MCMC to do inference, either
(1) sampling over the spherical
harmonic coefficients $\mathbf{y}$ and imposing a uniform prior in the range
$[0, 1]$ on the albedo values $\mathbf{a}$
evaluated on a discrete grid on the sphere or
(2) sampling over the albedo values $\mathbf{a}$ with the same uniform prior,
but using $\mathbf{y}$ to compute the actual light curve model. In both
cases, there exists a linear operator that transforms between
$\mathbf{y}$ and $\mathbf{a}$:
%
\begin{align}
    \label{eq:Py}
    \mathbf{a} = \mathbf{P} \mathbf{y}
\end{align}
%
and
%
\begin{align}
    \label{eq:PInvy}
    \mathbf{y} = \mathbf{P}^+ \mathbf{a}
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:PInv}
    \mathbf{P}^+ =
    (\mathbf{P}^\top \mathbf{P} + \lambda \mathbf{I})^{-1} \mathbf{P}^\top
\end{align}
%
is the pseudoinverse of $\mathbf{P}$, with (small) regularization parameter
$\lambda$ and where $\mathbf{I}$ is the identity matrix.
%
Each row of the matrix $\mathbf{P}$ is constructed from the value
of each of the spherical harmonics at the corresponding point on the
grid; both $\mathbf{P}$ and its inverse may be pre-computed for
efficiency. In both cases, the grid should be fine enough to ensure
positivity over most of the sphere but not so fine as to throttle the
computation; in practice, we find that grids with $2-5$ times as many pixels
as spherical harmonic coefficients are ideal.
%
The \href{https://rodluger.github.io/starry}{documentation}
includes tutorials on how to implement this in practice.

@@

\xxx{TODO.}

\vfill
\pagebreak

@@

\section{The Problem}
\label{sec:the-problem}
%
This paper closely follows the
notation and formalism introduced in \citet{Luger2019}. While we
include all of the relevant equations and definitions below,
the reader is encouraged to
review \citet{Luger2019} before proceeding.
To improve the readability of this paper,
Tables~\ref{tab:symbols}--\ref{tab:matrices} at the end list the principal
symbols and quantities used throughout the text, with links to the
sections and equations in which they are defined.

@@

An observer views the body from a large distance in the sky frame
$\mathcal{F}$, in which the $x$-axis points to the right,
the $y$-axis points up, and the $z$-axis points out of the sky
toward the observer. Following \citet{Luger2019}, if
an occultor of radius $r_o$ is located at sky position $(x_o, y_o)$,
we compute the visible flux $f$ from
%
\begin{align}
    \label{eq:sTARRy}
    f = \sTe \mathbf{A} \mathbf{R}' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, from right to left, $\mathbf{R} = \mathbf{R}(\text{I}, \Lambda, \Theta)$
is a Wigner rotation matrix that rotates $\bvec{y}$ from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$
given the body's inclination $\text{I}$, obliquity
$\Lambda$, and rotational phase $\Theta$
\citep[Appendix C in][]{Luger2019},
%
$\mathbf{R}' = \mathbf{R}'(x_o, y_o)$ rotates the body on the plane
of the sky into the integration frame $\mathcal{F}'$, in which the
occultor lies along the $+y'$-axis,
%
$\mathbf{A}$
\citep[Equation~B13 in][]{Luger2019}
is the change-of-basis matrix from $\by$
to the \emph{Green's basis} $\bg$ in which the integrals are computed,
whose component at index $n$ is
%
\begin{align}
    \label{eq:bg}
    \tilde{g}_{n}(x, y) & =
    \begin{dcases}
        %
        \frac{\mu+2}{2}x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        %
        z(x, y)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        3x^{l-2}yz(x, y)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \text{even}
        \\[1em]
        %
        z(x, y)
        \bigg(
        -x^{l-3} + x^{l-1} + 4x^{l-3}y^2
        \bigg)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \, \text{odd}
        \\[1em]
        %
        z(x, y)
        \bigg(
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu-1}{2}
        \ - \
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu+3}{2}
        \\
        \qquad\qquad \ - \
        \frac{\mu+3}{2} x^\frac{\mu-1}{2} y^\frac{\nu-1}{2}
        \bigg)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
with
%
\begin{align}
    \label{eq:mu-nu}
    \mu & \equiv l - m
    \nonumber          \\
    \nu & \equiv l + m
\end{align}
%
and
%
\begin{align}
    \label{eq:z}
    z(x, y) \equiv \sqrt{1 - x^2 - y^2}
    \quad,
\end{align}
%
and $\sTe = \sTe(b_o, r_o)$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bg$
\citep[Equation~26 in][]{Luger2019}, with $b_o = \sqrt{x_o^2 + y_o^2}$.

If instead no occultor is present, we compute the total
visible flux $f_0$ from this body as
%
\begin{align}
    \label{eq:rTA1Ry}
    f_0 = \rTe \mathbf{A_1} \mathbf{R}'' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, as before, $\mathbf{R} = \mathbf{R}(\text{I}, \Lambda, \Theta)$
rotates the body from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$,
%
$\mathbf{R}''$ rotates the body on the plane
of the sky into the integration frame
$\mathcal{F}''$%
\footnote{%
    In \citet{Luger2019}, $\mathcal{F}'' = \mathcal{F}'$, so
    this rotation is trivial: $\mathbf{R}''$ is just the identity matrix.
},
%
$\mathbf{A_1}$
\citep[Equation~B11 in][]{Luger2019}
is the change-of-basis matrix from the spherical harmonic
basis $\by$ to the \emph{polynomial basis} $\bp$ in which the integrals
are computed, whose component at index $n$ is
%
\begin{align}
    \label{eq:bp}
    \tilde{p}_n(x, y) & =
    \begin{dcases}
        x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        x^\frac{\mu-1}{2} y^\frac{\nu-1}{2} z(x, y)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
and $\rTe$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bp$
\citep[Equation~19 in][]{Luger2019}.

@@

For computational efficiency, in practice we compute the integral of the
intensity-weighted basis directly,
\xxx{NO LONGER TRUE. FIX THIS DEFINITION.}
    \label{eq:rTI}
    \overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}(b, r_s) & \equiv
    \rT(b) \mathbf{I}(b, \theta'', r_s)
    \mathbf{I}(b, 0, r_s)
recalling that in frame $\mathcal{F}''$, $\theta'' = 0$ by
definition.
Equation~(\ref{eq:rTI}) may be solved analytically in terms of purely

@@

The component of $\overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}$

@@

    \overset{\quad\wedge}{\mathbb{r}}_n(b, r_s) & =

@@

        \frac{b_c\left(1 - b^{\frac{\nu + 4}{2}}\right)}{2}
        \mathbb{L}_{\frac{\mu}{2}, \frac{\nu + 2}{2}} -
        b \mathbb{k}_\frac{\nu}{2}(b) \mathbb{M}_{\frac{\mu}{2}, \frac{\nu}{2}}

@@

        b_c

@@

        - \frac{b}{2} \bigg\{
        \left(1 - b^{\frac{\nu + 1}{2}}\right)
        \mathbb{L}_{\frac{\mu - 1}{2}, \frac{\nu - 1}{2}} -
        \mathbb{L}_{\frac{\mu + 3}{2}, \frac{\nu - 1}{2}}
        \qquad\qquad
        \left(1 - b^{\frac{\nu + 5}{2}}\right)
        \mathbb{L}_{\frac{\mu - 1}{2}, \frac{\nu + 3}{2}}
        \bigg\}

@@

Once $\overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}$ is known,
the observed flux is computed from
(c.f. Equation~\ref{eq:rTIA1RRy})

@@

    \overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}(b, r_s)

@@

    \hat{f}_0 & =
    -\overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}(-b, r_s)

@@

    \mathbf{R}''(-x_s, -y_s)

@@

\xxx{FIX THIS DEFINITION. NO LONGER HOW WE COMPUTE IT.}
This is the flux we would get if we placed a \emph{negative} illumination
source at the antipode $(-x_s, -y_s, -z_s)$ of the true source, and
will be useful in negating the unphysical negative contribution from our
polynomial illumination function (Equation~\ref{eq:illum_poly}) in
the following section.

>>>>

\keywords{methods: analytic --- techniques: photometric}

\begin{abstract}
    We derive efficient, closed form, differentiable, and numerically stable
    solutions for the flux measured from a planet or moon seen in reflected
    light, either in our out of occultation.
    %
    Our expressions apply to the computation of scattered
    light phase curves of exoplanets, secondary eclipse light curves in
    the optical, or future measurements of planet-moon and planet-planet
    occultations, as well as to photometry of solar system bodies.
    %
    We derive our solutions for Lambertian bodies illuminated by a point
    source, but extend them to model illumination sources of finite
    angular size and rough surfaces with phase-dependent scattering.
    %
    Our algorithm is implemented in \Python within the open-source
    \starry mapping framework and is designed with efficient gradient-based
    inference in mind.
    %
    The algorithm is ${\sim}4-5$ orders of magnitude faster than direct
    numerical evaluation methods and ${\sim}10$ orders of magnitude more
    precise.
    %
    We show how the techniques developed here may one day lead to the
    construction of two-dimensional maps of terrestrial planet surfaces,
    potentially enabling the detection of continents and oceans on
    exoplanets in the habitable zone.
    %
    \href{https://github.com/rodluger/starrynight}{\color{linkcolor}\faGithub}
\end{abstract}

%
\section{Introduction}
\label{sec:intro}
%

Despite recent advances in instrumentation and the dawn of
thirty meter-class telescopes and kilometer-wide interferomer arrays,
extrasolar planets will remain unresolved point sources for decades
to come. Nevertheless, modulations in the light received from these distant
bodies due to their rotation, changing illumination, and eclipses
by their host stars or other bodies in the system can be harnessed to
reconstruct two-dimensional views of their surfaces.
In particular, next-generation space-based telescopes such as the
Large UV/Optical/IR Surveyor (LUVOIR) may enable us to measure
variations in the reflected light signature of terrestrial planets in
the habitable zone, which can be used to map their surfaces and
indirectly infer the presence of clouds, continents, oceans, and perhaps
even life.

There is a vast literature on techniques for mapping exoplanet surfaces
based on their phase curves
\citep[e.g.,][]{Russell1906,Lacis1972,Knutson2007,Cowan2008,Oakley2009,Berdyugina2017}
and occultation light curves
\citep[e.g.,][]{Williams2006,Rauscher2007,Majeau2012,deWit2012,Rauscher2018},
both in thermal and reflected (scattered) light.
In particular, much attention has been given to techniques for mapping
Earth-like planets from visible-light reflected phase curves
\citep[e.g.,][]{Ford2001,Kawahara2010,Kawahara2011,Fujii2012,Kawahara2020,Aizawa2020}.
Unlike thermal phase curves, which primarily encode (often degenerate)
information about longitudinal surface brightness variations
\citep{Russell1906}, reflected light curves often contain information
about the full two-dimensional surface albedo distribution
\citep[e.g.,][]{Kawahara2010}.

Occultation light curves in reflected light can encode even more information
about the surface. Thus far, these have been studied primarily within our
solar system. Mutual occultations among the Galilean moons of Jupiter
have been studied to infer surface properties of Io \xxx{...more on this,
    and add citations.}

Many studies have analyzed real Earth reflected light curves to
infer surface properties of our planet as an exercise in
prepartion for the mapping of exoplanets.
%
\citet{Cowan2009} analyzed visible-light disk-integrated light curves
of the Earth taken by the Deep Impact spacecraft to produce
longitudinal maps of the surface, harnessing multi-band observations
to disentangle static surface brightness features from temporally
variable clouds. More recently, \citet{Jiang2018} and \citet{Fan2019}
used data from the L1-stationed DSCOVR satellite to infer surface
and cloud properties of the Earth, and \citet{Luger2019b} analyzed
background scattered light in TESS photometry to reconstruct a cloud
map of the Earth.

On the open-source software front, \citet{Louden2018} developed
\textsf{spiderman}, an efficient discretization scheme
on the sphere that enables fast computation of exoplanet
phase curves and occultation light curves in \Python.
%
\citet{Haggard2018} presented
\textsf{EARL} (Exoplanet Analytic Reflected Lightcurves),
a \textsf{Mathematica} code to compute analytic, closed form
solutions for the phase curve of a Lambert sphere, i.e., one that
scatters light isotropically, in the case that the surface albedo
distribution is characterized by either a sum of delta functions
or a sum of spherical harmonics.
%
\citet{Farr2018} released \textsf{exocartographer}, a Bayesian
framework for doing inference on exoplanet phase curves based on
a HEALPix \citep{healpix} discretization scheme.
%
Finally, \citet{Luger2019} introduced \starry, a light curve
modeling package that computes thermal phase curves and occultation
light curves, as well as their derivatives, analytically from a spherical
harmonic expansion of the surface brightness.

The present paper is an extension to the \starry algorithm,
adapting it to model phase curves and occultation light curves
in reflected light. The expressions
we derive are analytic: they may all be expressed in closed form
in terms of algebraic operations involving trigonometric functions
and (at times) elliptic integrals. We derive numerically
stable recursion relations for the efficient evaluation of all
expressions and code them within an autodifferentiation framework
to enable the computation of high precision derivatives for use in
gradient-based inference and optimization schemes.
Our code is fully
\href{https://github.com/rodluger/starry}{open-source},
comprehensively
\href{https://dev.azure.com/rodluger/starry/_test/analytics?definitionId=4}{unit-tested},
and supplemmented with an extensive
\href{https://rodluger.github.io/starry}{API documentation}
and suite of
\href{https://rodluger.github.io/starry/latest/tutorials.html}{tutorials}.
%
As in \citet{Luger2019}, in the caption of each of the figures
we provide links \codeicon\ to the exact \Python scripts that
generated them. Next to many of the equations we also provide
links \prooficon\ to \textsf{Jupyter} notebooks containing
detailed derivations and/or validations.

\section{Overview}
\label{sec:overview}

@@

The piecewise nature of the illumination function at the day/night terminator
makes the problem of computing the visible flux particularly difficult; most
studies to date have tackled the problem numerically, either
via Monte Carlo integration \citep[e.g.,][]{Ford2001} or by discretizing the surface
and computing the relevant integrals by summing over the visible
pixels \xxx{(cite)}. Recently, \citet{Haggard2018} developed an analytic
framework for computing light curves of unocculted bodies illuminated
by a point source. In this paper, we re-derive their solution under the
\starry framework and extend it for the first time to the case where
the body is occulted by another spherical body, which may or may not be the
illumination source. We also extend the solution to the case of an extended
illumination source and to non-Lambertian scattering.

@@

This paper is organized as follows.
In \S\ref{sec:validation} we present sample light curves computed
using our algorithm, validate it against numerical integration, and
discuss its performance in terms of computational speed and precision.
In \S\ref{sec:extensions} we extend the model to apply to illumination
sources of finite size and surfaces that scatter light anisotropically.
We discuss implications, applications, and limitations of our model in
\S\ref{sec:discussion} and summarize our findings in \S\ref{sec:conclusions}.
For convenience, Tables~\ref{tab:symbols}--\ref{tab:matrices} at the end
list all symbols and variables used in the text, with descriptions and
links to the equations in which they are defined.

@@

The algorithm presented in this paper has been implemented in the \Python
package \starry, which can be installed from
\href{https://github.com/rodluger/starry}{GitHub} or via the Python package
manager \textsf{pip}. The algorithm is coded in a mixture of \textsf{C++}
with forward automatic differentiation using the \textsf{Eigen} library
\citep{eigen} and \Python with backward differentiation using
just-in-time compiled \textsf{theano}
operations \citep{theano}. The user interface, however, is purely in
\Python. The \textsf{theano} backend facilitates integration with the
\textsf{exoplanet} modeling package \citep{exoplanet} and in particular
with \textsf{pymc3} \citep{pymc3} for inference with gradient-based
Markov Chain Monte Carlo (MCMC) schemes
such as Hamiltonian Monte Carlo \citep[HMC;][]{Duane1987}
and No-U-Turn Sampling \citep[NUTS;][]{Hoffman2011}.
%
Complete \href{https://rodluger.github.io/starry}{documentation} and an
extensive library of tutorials is available online.
The links next to each
of the figures \codeicon\, point to the \Python scripts used to generate
them and may also help in learning how to use \starry.

@@

The third limitation concerns the flexibility of the \starry model.
While we presented ways to capture non-Lambertian scattering
in \starry (\S\ref{sec:nonlambertian}), there are certain aspects
of light curves in reflected light that cannot be captured by the model.
One example of this is shadowing. Craters on the moon or
volcanoes on Io can cast large shadows visible from space, particularly
if viewed near crescent phase. Unfortunately, there is no way to model
this within the \starry framework. Another example is multiple scattering,
as in optically thick atmospheres, for which a proper radiative transfer model
must be used. There is also the case of specular reflection,
or ``glint'', which is a pronounced signal for the Earth due its
oceans \citep{Robinson2014} and on Titan due to
its hydrocarbon lakes \citep{Barnes2011}. In principle, glint could be
modeled in the same way as non-Lambertian sccattering, by constructing the
linear operator $\mathbf{I}$ in such a way as to downweight portions of the
projected disk where $\vartheta_i \neq \vartheta_r$. In practice, however,
if the size of the glint spot is small (which is typically the case),
an expansion at extremely high $l$ ($l \sim 360$ in the case of the Earth)
would be required, which would not work for the reasons above.
Instead, it may be possible to combine the \starry algorithm with the
formalism of \citet{Haggard2018}, who derived analytic expressions for
phase curves of delta function maps ($\delta$-map), to model glint.
As discussed in \citet{LustigYaeger2018}, glint mapping is an extremely
powerful way to not only map terrestrial planets but also to confirm
their habitability via the presence of an ocean.

Finally, we would like to emphasize that while spherical harmonics are
a convenient basis for the purpose of computing light curves, they
have a significant drawback when it comes to modeling real planetary
surfaces: it can be difficult to strictly enforce physical
values of the albedo everywhere on the surface map when doing inference.
That is because there is no analytic way to determine whether a
spherical harmonic representation is positive-valued
(or restricted to a given range) everywhere on the sphere.
Instead, this must be checked numerically, by evaluating the function on a
discrete grid. This makes it somewhat cumbersome to implement positivity as a prior
when doing inference; in particular, this prior cannot be expressed as
a Gaussian, so the analytic expression for the posterior
discussed in \S\ref{sec:linearity} will generally have nonzero support for
negative albedo values. This is particularly problematic when the data is
not very constraining and the posterior is prior-dominated, as
positivity can be an \emph{extremely} informative prior
\citep[e.g.,][]{Fienup1978}.
We therefore recommend that in such cases
users of the \starry algorithm use MCMC to do inference, either
(1) sampling over the spherical
harmonic coefficients $\mathbf{y}$ and imposing a uniform prior in the range
$[0, 1]$ on the albedo values $\mathbf{a}$
evaluated on a discrete grid on the sphere or
(2) sampling over the albedo values $\mathbf{a}$ with the same uniform prior,
but using $\mathbf{y}$ to compute the actual light curve model. In both
cases, there exists a linear operator that transforms between
$\mathbf{y}$ and $\mathbf{a}$:
%
\begin{align}
    \label{eq:Py}
    \mathbf{a} = \mathbf{P} \mathbf{y}
\end{align}
%
and
%
\begin{align}
    \label{eq:PInvy}
    \mathbf{y} = \mathbf{P}^+ \mathbf{a}
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:PInv}
    \mathbf{P}^+ =
    (\mathbf{P}^\top \mathbf{P} + \lambda \boldsymbol{\mathsf{I}})^{-1} \mathbf{P}^\top
\end{align}
%
is the pseudoinverse of $\mathbf{P}$, with (small) regularization parameter
$\lambda$ and where $\boldsymbol{\mathsf{I}}$ is the identity matrix.
%
Each row of the matrix $\mathbf{P}$ is constructed from the value
of each of the spherical harmonics at the corresponding point on the
grid; both $\mathbf{P}$ and its inverse may be pre-computed for
efficiency. In both cases, the grid should be fine enough to ensure
positivity over most of the sphere but not so fine as to throttle the
computation; as a rule of thumb, we find that grids with $2-5$ times
as many pixels
as spherical harmonic coefficients are sufficient.
%
The \href{https://rodluger.github.io/starry}{documentation}
includes tutorials on how to implement this in practice.

@@

We have presented an efficient, numerically stable,
closed-form algorithm for computing phase curves
and occultation light curves of spherical bodies in reflected
(scattered) light.
This algorithm is an extension of the algorithm presented
in \citet{Luger2019} for light curves in thermal light and
is generally applicable to exoplanetary phase
curves, secondary eclipses, and occultations by moons and other planets,
as well as to light curves of planets and moons in our solar system.
We derive the solution for the case of a Lambert sphere illuminated by a
point source, but extend it to the case of an extended illumination
source and non-Lambertian scattering parametrized by a surface roughness
coefficient.
%
The algorithm is ${\sim}4-5$ orders of magnitude faster
and ${\sim}10$ orders of magnitude more precise than other numerical
approaches for computing these light curves. The algorithm is
also differentiable, enabling inference with efficient gradient-based samplers
such as Hamiltonian Monte Carlo (HMC), and linear in the spherical
harmonic coefficients describing the surface albedo, enabling fast,
closed-form solutions for the albedo posterior distribution under a
Gaussian noise model.
%
We implement the algorithm within the \starry software, an open-source \Python
package for inferring surface maps of unresolved celestial bodies. The
algorithm is coded in a combination of \cpp and \Python compiled using
the \textsf{theano}. The interface was designed specifically for
compatibility with the \textsf{exoplanet} modeling package
\citep{exoplanet} and the \textsf{pymc3} inference suite \citep{pymc3}.

Upcoming telescopes will enable measurements of exoplanet phase curves
and secondary eclipses at unprecendented precision. While the
James Webb Space Telescope (JWST) will be primarily sensitive to thermal
emission from exoplanets (which can currently be modeled with \starry),
next-generation direct imaging facilities such as the
Large UV/Optical/IR Surveyor (LUVOIR)
will enable measurements in reflected light, in particular for terrestrial
planets in the habitable zone. Because of the changing illumination
pattern over the course of an orbit of the planet, phase curves and
occultation light curves in reflected
light contain vastly more information about the two-dimension albedo
distribution of the body than their thermal counterparts.
With careful modeling, light curves in reflected light are likely to give us
the first images of potentially habitable exoplanets, enabling the
detection of clouds, continents, oceans, and perhaps even life.

\vspace{1em}

The software presented in this work is open source under the MIT License and
is available at \url{https://github.com/rodluger/starry}, with documentation
and tutorials hosted at \url{https://rodluger.github.io/starry}.
A permanent version of the code used to generate the figures and results in
this paper is archived at \xxx{https://doi.org/XX.XXXX/zenodo.XXXXXXX}.

\software{
    Eigen v3 \citep{eigen},
    exoplanet \citep{exoplanet},
    HEALPix \citep{healpix},
    pybind11 \citep{pybind11},
    pymc3 \citep{pymc3},
    scipy \citep{scipy},
    starry \citep{starry}.
}

@@

\section{The Problem}
\label{sec:the-problem}
%
This paper closely follows the
notation and formalism introduced in \citet{Luger2019}. While we
include all of the relevant equations and definitions below,
the reader is encouraged to
review \citet{Luger2019} before proceeding.
To improve the readability of this paper,
Tables~\ref{tab:symbols}--\ref{tab:matrices} at the end list the principal
symbols and quantities used throughout the text, with links to the
sections and equations in which they are defined.
Because of the large number of symbols used in this paper, we adopt
the following conventions: scalars are represented by regular lowercase or
occasionally uppercase letters (i.e., $x$ or $X$), vectors are
represented by boldface lowercase letters
($\mathbf{x}$), and matrices and other linear operators are represented
by boldface capital letters ($\mathbf{X}$).
With a few exceptions,
Greek letters are reserved for angular quantities and may be either
scalars ($\alpha$) or vectors ($\pmb{\alpha}$).
Script font is typically used to denote curves or frames of reference
($\mathcal{F}$). Primes are used to
distinguish between frames of reference ($x$ and $x'$ are used
to denote the same quantity, but in frames $\mathcal{F}$ and $\mathcal{F}'$,
respectively). Tildes are used to denote basis vectors ($\by$). Finally,
blackboard vectors ($\mathbbb{x}$) correspond to solutions
to the various ``primitive'' integrals that arise in the occultation
problem.

@@

An observer views the body from a large distance in the sky frame
$\mathcal{F}$, in which the $x$-axis points to the right,
the $y$-axis points up, and the $z$-axis points out of the sky
toward the observer. Following \citet{Luger2019}, if
an occultor of radius $r_o$ is located at sky position $(x_o, y_o)$,
we compute the visible thermal flux $f_T$ from
%
\begin{align}
    \label{eq:sTARRy}
    f_T = \sTe \mathbf{A} \mathbf{R}' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, from right to left, $\mathbf{R} = \mathbf{R}(\text{I}, \Lambda, \Theta)$
is a Wigner rotation matrix that rotates $\bvec{y}$ from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$
given the body's inclination $\text{I}$, obliquity
$\Lambda$, and rotational phase $\Theta$
\citep[Appendix C in][]{Luger2019},
%
$\mathbf{R}' = \mathbf{R}'(x_o, y_o)$ rotates the body on the plane
of the sky into the integration frame $\mathcal{F}'$, in which the
occultor lies along the $+y'$-axis,
%
$\mathbf{A}$
\citep[Equation~B13 in][]{Luger2019}
is the change-of-basis matrix from $\by$
to the \emph{Green's basis} $\bg$ in which the integrals are computed,
whose component at index $n$ is
%
\begin{align}
    \label{eq:bg}
    \tilde{g}_{n}(x, y) & =
    \begin{dcases}
        %
        \frac{\mu+2}{2}x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        %
        z(x, y)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        3x^{l-2}yz(x, y)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \text{even}
        \\[1em]
        %
        z(x, y)
        \bigg(
        -x^{l-3} + x^{l-1} + 4x^{l-3}y^2
        \bigg)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \, \text{odd}
        \\[1em]
        %
        z(x, y)
        \bigg(
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu-1}{2}
        \ - \
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu+3}{2}
        \\
        \qquad\qquad \ - \
        \frac{\mu+3}{2} x^\frac{\mu-1}{2} y^\frac{\nu-1}{2}
        \bigg)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
with
%
\begin{align}
    \label{eq:mu-nu}
    \mu & \equiv l - m
    \nonumber          \\
    \nu & \equiv l + m
\end{align}
%
and
%
\begin{align}
    \label{eq:z}
    z(x, y) \equiv \sqrt{1 - x^2 - y^2}
    \quad,
\end{align}
%
and $\sTe = \sTe(b_o, r_o)$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bg$
\citep[Equation~26 in][]{Luger2019}, with $b_o = \sqrt{x_o^2 + y_o^2}$.

If instead no occultor is present, we compute the total
visible thermal flux $f_{T_0}$ from this body as
%
\begin{align}
    \label{eq:rTA1Ry}
    f_{T_0} = \rTe \mathbf{A_1} \mathbf{R}'' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, as before, $\mathbf{R} = \mathbf{R}(\text{I}, \Lambda, \Theta)$
rotates the body from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$,
%
$\mathbf{R}''$ rotates the body on the plane
of the sky into the integration frame
$\mathcal{F}''$%
\footnote{%
    In \citet{Luger2019}, $\mathcal{F}'' = \mathcal{F}'$, so
    this rotation is trivial: $\mathbf{R}''$ is just the identity matrix.
},
%
$\mathbf{A_1}$
\citep[Equation~B11 in][]{Luger2019}
is the change-of-basis matrix from the spherical harmonic
basis $\by$ to the \emph{polynomial basis} $\bp$ in which the integrals
are computed, whose component at index $n$ is
%
\begin{align}
    \label{eq:bp}
    \tilde{p}_n(x, y) & =
    \begin{dcases}
        x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        x^\frac{\mu-1}{2} y^\frac{\nu-1}{2} z(x, y)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
and $\rTe$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bp$
\citep[Equation~19 in][]{Luger2019}.

@@

Equation~(\ref{eq:rT}) may be solved analytically in terms of purely

@@

The component of $\rT$

@@

    \mathbb{r}_n(b) & =

@@

        \frac{\left(1 - b^{\frac{\nu + 4}{2}}\right)}{2}
        \mathbb{L}_{\frac{\mu}{2}, \frac{\nu + 2}{2}}

@@

@@

@@

Once $\rT$ is known,
the observed total flux in reflected light
is computed from (c.f. Equation~\ref{eq:rTIA1RRy})

@@

    \rT(b)
    \mathbf{I}(b, \theta'', r_s)

@@

    \hat{f}_0 &=
    \Big(
    \mathbf{r}^\top
    \rT(b)
    \Big)
    \mathbf{I}(b, \theta'', r_s)

@@

@@

This is the flux contribution from the unphysical night side
(if we were to integrate over it), where
our polynomial illumination (Equation~\ref{eq:illum_poly}) function
yields \emph{negative} intensities. This quantity
will be useful in negating the unphysical contribution
in the integrals of the following section.
